---
title: "Time Series Investigations"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE}

rm(list = ls())
source("header.R")

df <- st_read("data/mst/hex_merged_w_stats.gpkg")

df <- df %>% 
  mutate(bk_res_4 = bk_res_4_1 + bk_res_4_2 + bk_res_4_3, 
         bk_res_5 = bk_res_5_1 + bk_res_5_2 + bk_res_5_3) %>% 
  mutate(bk_res_4_cat = case_when(bk_res_4_3 > 0 ~ "Any Increase From 1% or More to More than 90%", 
                                  bk_res_4_2 > 0 ~ "Significant Increase From 1% or More to Less than 90%", 
                                  bk_res_4_1 > 0 ~ "Any Increase From Less Than 1%")) %>% 
  mutate(bk_res_5_cat = case_when(bk_res_5_3 > 0 ~ "Any Increase From 1% or More to More than 90%", 
                                  bk_res_5_2 > 0 ~ "Significant Increase From 1% or More to Less than 90%", 
                                  bk_res_5_1 > 0 ~ "Any Increase From Less Than 1%")) %>% 
  mutate(bk_res_since_39 = ifelse((bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & bk_res_4 == 0, 
                                  "Black Residency Since 1939", "No Black Residency in 1950"))




```

# Plot of Homicide Rate Over Time


```{r, echo=FALSE}

years <- seq(1880, 1965, 5)

year_select <- years[1]

df %>% st_drop_geometry() %>% filter(hom_rt_mean_5>0) %>% 
  summarise(
            q_99_rt = quantile(hom_rt_mean_5, 0.99, na.rm = T), 
            q_999_rt = quantile(hom_rt_mean_5, 0.999, na.rm = T), 
            q_99_ct = quantile(hom_ct_sum_5, 0.99, na.rm = T), 
            q_999_ct = quantile(hom_ct_sum_5, 0.999, na.rm = T), 
            max = max(hom_ct_sum_5, na.rm = T)) 

for (year_select in years) {
  


  g <- ggplot(data = df %>% filter(year == year_select) %>% 
                mutate(hom_rt_mean_5 = ifelse(is.na(hom_rt_mean_5), 0, hom_rt_mean_5))) +
    geom_sf(aes(fill = hom_rt_mean_5, color = as.factor(in_check))) +
    theme_minimal() + 
    labs(title = glue("Homicide Rate {year_select-5}-{year_select}"), 
         fill = "Average Homicide Rate", color = "In Checkerboard") + 
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 3), 
      na.value = "white") 
  
  plot(g)

  
}

for (year_select in years) {
  
  g <- ggplot(data = df %>% filter(year == year_select)) +
    geom_sf(aes(fill = hom_ct_sum_5, color = as.factor(in_check))) +
    theme_minimal() +
    labs(title = glue("Number of Homicides {year_select-5}-{year_select}"), 
         fill = "Total Number of Homicides",  color = "In Checkerboard") +
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 80), 
      na.value = "white") 


  plot(g)

  
}
 
```

# Preliminary Identification of Changes 

Calculate the change of total homicides over 5 year periods, the change of average homicides over 5 year periods, and the change of average homicide rate over 5 year periods. 

```{r, echo=FALSE}

df_changes <- df %>% 
  st_drop_geometry() %>% 
  #remove years with lack of data
  filter(year_5 != 1880 & year_5 != 1930 & year_5 != 1935 & year_5 != 1940 & year_5 != 1945 ) %>% 
  #filter to period of interest 
  filter(year_5 <= 1965) %>%
  distinct(GRID_ID, year_5, hom_ct_sum_5, hom_ct_mean_5, hom_rt_mean_5) %>% 
  arrange(year_5) %>% 
  group_by(GRID_ID) %>% 
  mutate(across(c(hom_ct_sum_5, hom_ct_mean_5, hom_rt_mean_5), ~(. - lag(.)))) %>% 
  filter(year_5 > 1875) %>% 
  ungroup() 



change_calc <- function(var, df = df_changes, decline = TRUE) {

  df["change"] <- df[var]
  
  df <- df %>% 
      distinct(GRID_ID, year_5, change) 
  
  if (decline) {
    df <- df %>%  
      filter(change < 0) %>% 
      arrange(change) %>% 
      group_by(year_5) %>% 
      mutate(pctile = percent_rank(desc(change))) %>%
      mutate(type = glue("decline_{var}")) 
  } else {
    df <- df %>%
      filter(change > 0) %>% 
      arrange(desc(change))%>% 
      group_by(year_5)%>% 
      mutate(pctile = percent_rank(change)) %>%
      mutate(type = glue("rise_{var}")) 
    
  }
  
  df <- df %>% 
    ungroup() %>% 
    filter(pctile > 0.9) 
  
  return(df)
  
}

decline_sum_ct <- change_calc("hom_ct_sum_5", decline = T)

decline_mean_ct <- change_calc("hom_ct_mean_5", decline = T)

decline_mean_rt <- change_calc( "hom_rt_mean_5", decline = T)
  

rise_sum_ct <- change_calc("hom_ct_sum_5", decline = F)

rise_mean_ct <- change_calc("hom_ct_mean_5", decline = F)

rise_mean_rt <- change_calc("hom_rt_mean_5", decline = F)


df_changes_sum <- rbind(decline_sum_ct, decline_mean_ct, decline_mean_rt, rise_sum_ct, rise_mean_ct, rise_mean_rt)

```


## Notable Events 

### 1950s - 1960s

```{r, echo=FALSE}

year_c <- 1955
var <- "hom_ct_sum_5"

plot_rise <- function(year_c, var) {

  #Get riots in that time period
  riots <- df %>% 
    st_drop_geometry() %>% 
    filter(!is.na(name_riot)) %>% distinct(GRID_ID, year) %>% 
    filter(year >= year_c-5 & year <= year_c) %>% 
    pull(GRID_ID)
  
  #Get significant increases in that time period
  IDS_rise <- df_changes_sum %>% 
    filter(type == glue("rise_{var}"),year_5 == year_c) %>% 
    distinct(GRID_ID) %>% 
    pull()
  
  IDS_decline <- df_changes_sum %>% 
    filter(type == glue("decline_{var}"),year_5 == year_c) %>% 
    distinct(GRID_ID) %>% 
    pull()
  
  bk_res_change_4 <- df %>% 
    filter(year == year_c) %>%
    filter(!is.na(bk_res_4_cat)) %>% 
    pull(GRID_ID)
  
  #Filter data to those IDS
  
  df_plot <- df %>% 
    filter(year == year_c) %>%
    mutate(flag_rise = ifelse(GRID_ID %in% IDS_rise,  "Notable Rise", NA), 
           flag_decline = ifelse(GRID_ID %in% IDS_decline , "Notable Decline", NA), 
         flag_riot = ifelse(GRID_ID %in% riots , "Riot", NA), 
         flag_res_change =  ifelse(GRID_ID %in% bk_res_change_4 , 
                                   "Black Migration 1939-1950", bk_res_since_39)) %>% 
    mutate(flag = paste0(flag_rise, ifelse(is.na(flag_rise), "", ifelse(!is.na(flag_decline) | 
                                                                          !is.na(flag_riot) | 
                                                                          !is.na(flag_res_change)," & ", "")),
                        flag_decline, ifelse(is.na(flag_decline), "", ifelse(!is.na(flag_riot) | 
                                                                          !is.na(flag_res_change)," & ", "")), 
                         flag_riot, ifelse(is.na(flag_riot), "", ifelse(!is.na(flag_res_change), " & ", "")), 
                        flag_res_change), 
           flag = str_replace_all(flag, "NA", "")) %>% 
    filter(!is.na(flag)) %>% 
    mutate(bk_res_4_cat = ifelse(is.na(bk_res_4_cat), "No Change", bk_res_4_cat))
  
  
    #Labels
  
   plot_colors <- setNames(c("#c39797", "#632523","red",
                              "#ffc04c", "#ffdfbf",
                             
                            "#A8DCAB", "#80EF80", 
                            "#00FF00",
                            "#05472A", "#228B22",
                            
                            "#6897bb", "#254061", "#D3D3D3", 
                            
                            "#f23fb2", 
                            "purple", "#e66fa0",
                            
                            "black", "black",  
                            "black", "black", 
                             "black", "black"),
                          c("Notable Rise", "Notable Rise & Black Migration 1939-1950", "Notable Rise & Riot",
                          "Notable Rise & Black Residency Since 1939", "Notable Rise & No Black Residency in 1950", 
                          
                          "Notable Decline", "Notable Decline & Black Migration 1939-1950",  
                          "Notable Decline & Riot", 
                          
                          "Notable Decline & Black Residency Since 1939", 
                          "Notable Decline & No Black Residency in 1950", 
                          
                          "Black Migration 1939-1950", 
                          "Black Residency Since 1939", "No Black Residency in 1950", 
                          
                          "Riot",  
                          "Riot & No Black Residency in 1950", "Riot & Black Residency Since 1939",
                          
                          "Notable Rise & Riot & Black Migration 1939-1950",  
                          "Notable Rise & Riot & Black Residency Since 1939", 
                          "Notable Rise & Riot & No Black Residency in 1950",
                          
                          "Notable Decline & Riot & Black Migration 1939-1950",
                          "Notable Decline & Riot & Black Residency Since 1939", 
                          "Notable Decline & Riot & No Black Residency in 1950")) 
   
  
  plot_colors <- plot_colors[df_plot %>% distinct(flag) %>% pull()]
  
  plot_colors_lines <- setNames(c( "#e5ff4c", "#ff7f00","#836FFF",
                                  "white"),
                                c("Any Increase From 1% or More to More than 90%",
                                  "Significant Increase From 1% or More to Less than 90%",
                                  "Any Increase From Less Than 1%", "No Change"))


  plot_colors_lines <- plot_colors_lines[df_plot %>% distinct(bk_res_4_cat)  %>% pull()]

  
    #Plot
  g <- ggplot(data = df_plot) +
      geom_sf(aes(fill = flag, color = as.factor(in_check))) +
      theme_minimal() + 
      labs(title =  glue("Significant Events {year_c}"))+ 
      scale_fill_manual(values = plot_colors) +
      scale_color_manual(values = c("darkgrey", "black")) 
    
  plot(g)
  
  #Plot
  g <- ggplot(data = df_plot) +
      geom_sf(aes(fill = flag, color = bk_res_4_cat), linewidth = 0.55) +
      theme_minimal() + 
      labs(title =  glue("Significant Events {year_c}"))+ 
      scale_fill_manual(values = plot_colors) +
      scale_color_manual(values = plot_colors_lines) 
    
  plot(g)
  
  df_plot %>% 
    st_drop_geometry() %>% 
    group_by(flag) %>% 
    summarise(count = n()) %>% 
    ungroup() %>% 
    arrange(flag)
  
}

plot_rise(1950, "hom_ct_sum_5")

plot_rise(1955, "hom_ct_sum_5")

plot_rise(1960, "hom_ct_sum_5")

plot_rise(1965, "hom_ct_sum_5")



plot_rise(1950, "hom_ct_mean_5")

plot_rise(1955, "hom_ct_mean_5")

plot_rise(1960, "hom_ct_mean_5")

plot_rise(1965, "hom_ct_sum_5")


plot_rise(1950, "hom_rt_mean_5")

plot_rise(1955, "hom_rt_mean_5")

plot_rise(1960, "hom_rt_mean_5")

plot_rise(1965, "hom_ct_sum_5")

```





## Investigate Riot Years 



## Investigate Zoning 


## Investigate Change in Black Residency to 1950

```{r, echo=FALSE}



df_45_50_change <- df %>% filter(year %in% c(1950, 1945)) %>% 
  dplyr::select(-year_5) %>%
  dplyr::select(GRID_ID, year, contains("sum_5"), contains("mean_5")) %>% 
  arrange(year) %>% 
  group_by(GRID_ID) %>% 
  mutate(across(contains("_5"), .fns = list(change = ~(.-lag(.))))) %>% 
  ungroup() %>%
  filter(year == 1950)

# test <- df %>% 
#   st_drop_geometry()  %>% 
#   filter(bk_res_4 > 0 | bk_res_5> 0) %>%
#   dplyr::select(GRID_ID, contains("res_")) %>% 
#   distinct() 

ggplot(data = df %>% 
                filter(year == 1950) %>% 
                filter(bk_res_4 >0)) +
    geom_sf(aes(fill = bk_res_4_cat)) +
    theme_minimal() + 
    labs(title = "Increase 1939 to 1950")+ 
    coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 
  
  

ggplot(data = df %>% 
                filter(year == 1950) %>% 
                filter(bk_res_5 >0)) +
    geom_sf(aes(fill = bk_res_5_cat)) +
    theme_minimal() + 
    labs(title =  "Increase 1948 to 1950")+ 
    coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 
  

ggplot(data = df_45_50_change %>% filter(hom_ct_sum_5_change > 0)) +
    geom_sf(aes(fill = hom_ct_sum_5_change)) +
    theme_minimal() + 
    labs(title =  "Absolute Increase 1945 to 1950")+ 
    coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 

ggplot(data = df_45_50_change %>% filter(hom_rt_mean_5_change > 0)) +
    geom_sf(aes(fill = hom_rt_mean_5_change)) +
    theme_minimal() + 
    labs(title =  "Rate Increase 1945 to 1950")+ 
    coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 

ggplot(data = df_45_50_change %>% filter(hom_ct_sum_5_change < 0)) +
    geom_sf(aes(fill = hom_ct_sum_5_change)) +
    theme_minimal() + 
    labs(title =  "Absolute Decrease 1945 to 1950")+ 
    coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 

ggplot(data = df_45_50_change %>% filter(hom_rt_mean_5_change < 0)) +
    geom_sf(aes(fill = hom_rt_mean_5_change)) +
    theme_minimal() + 
    labs(title =  "Rate Decrease 1945 to 1950")+ 
    coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 
  

```



# Correlations 


## Rough Correlation Check 

```{r, echo = FALSE}
df_no_geom <- df %>% st_drop_geometry()

calc_cor <- function(var, control_var = "hom_ct_sum_5", data = df_no_geom) {
  print(var)
 correlation <-  cor(data[control_var], data[var], use = "complete.obs")[1]
 output <- data.frame(variable = var, 
                      correlation = correlation)
 return(output)
}

cols <- c("year_riot", "in_check", "still_desi",  "def_decl", "hazardous" , "industrial", "best", "commercial" , 
          "bk_res_1" , "bk_res_2", "bk_res_3", "bk_res_4_1", "bk_res_4_2" , "bk_res_4_3" , "bk_res_5_1" ,  "bk_res_5_2", 
          "bk_res_5_3" )
output <- map_df(cols, calc_cor)

output %>% 
  arrange(-abs(correlation))

```

# Testing 


## Testing Black Residency Values


We see polygons where the percentage of 1% concentration is lower than 10% or 90%. There are very few instances where the difference is large (more than 10 percentage points difference in area mapped), and most occur in the south of chicago where the maps were a) most distorted and b) the least amount of development matches the present state. 

It may be best to use the 90% value rather than the 10% or 1% value in these instances. This is further corroborated by the original maps. The areas plotted below where the difference is highest appear to be of fairly equal size in all three maps. 


```{r}


test <- df %>%
  filter(year == 1950) %>% 
  filter(bk_res_3 < bk_res_2 | bk_res_2 < bk_res_1 | bk_res_3 < bk_res_1) %>%
  dplyr::select(GRID_ID, bk_res_1, bk_res_2, bk_res_3) %>% 
  mutate(diff_1_2 = bk_res_1 - bk_res_2, 
         diff_1_3 = bk_res_1 - bk_res_3, 
         diff_2_3 = bk_res_2 - bk_res_3) %>% 
  pivot_longer(cols = contains("diff")) %>% 
  filter(value > 0.1) 

test 

ggplot(data = test ) + 
  geom_sf(aes(fill = GRID_ID)) + 
  coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 
  
```
