---
title: "Time Series Investigations"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE}

rm(list = ls())
source("header.R")

df <- st_read("data/mst/hex_merged_w_stats.gpkg")



```


### Estimation of Black Residency in 1939 and 1948 
```{r, echo=FALSE}

df_black_res <- df %>% 
  st_drop_geometry() %>% 
  filter(year == 1950) %>% 
  dplyr::select(GRID_ID, contains("bk_res")) %>% 
  mutate(est_black_res_39 = pmax(bk_res_1, bk_res_2, bk_res_3) - (bk_res_4_1 + bk_res_4_2 + bk_res_4_3), 
         est_black_res_39 = ifelse(est_black_res_39 <=  0, 0, est_black_res_39)) %>% 
    mutate(est_black_res_48 = abs(pmax(bk_res_1, bk_res_2, bk_res_3) - (bk_res_5_1 + bk_res_5_2 + bk_res_5_3)), 
         est_black_res_48 = ifelse(est_black_res_48 <=0, 0, est_black_res_48)) %>% 
  dplyr::select(GRID_ID, est_black_res_39, est_black_res_48)

df <- df %>% 
  left_join(df_black_res, by = "GRID_ID") %>% 
  mutate(bk_res_4 = bk_res_4_1 + bk_res_4_2 + bk_res_4_3, 
         bk_res_5 = bk_res_5_1 + bk_res_5_2 + bk_res_5_3) %>% 
  mutate(bk_res_4_cat = case_when(bk_res_4_2 > 0 ~ "Significant Increase From 1% or More to Less than 90%", 
                                  bk_res_4_3 > 0 ~ "Any Increase From 1% or More to More than 90%", 
                                  bk_res_4_1 > 0 ~ "Any Increase From Less Than 1%")) %>% 
  mutate(bk_res_since_39 = case_when((bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & est_black_res_39 > 0 & bk_res_4 == 0 ~ 
                                       "Black Residency Likely Since 1939 & No Migration", 
                                     (bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & est_black_res_39 > 0 & bk_res_4 > 0 ~ 
                                       "Black Residency Likely Since 1939 & Migration",
                                      (bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & est_black_res_39 == 0 & bk_res_4 > 0 ~ 
                                       "Likely No Black Residency in 1939 & Migration",
                                     TRUE ~ "No Black Residency in 1950")) %>% 
    mutate(bk_res_5_cat = case_when(bk_res_5_2 > 0 ~ "Significant Increase From 1% or More to Less than 90%", 
                                  bk_res_5_3 > 0 ~ "Any Increase From 1% or More to More than 90%", 
                                  bk_res_5_1 > 0 ~ "Any Increase From Less Than 1%")) %>% 
  mutate(bk_res_since_48 = case_when((bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & est_black_res_48 > 0 & bk_res_5 == 0 ~ 
                                       "Black Residency Likely Since 1948 & No Migration", 
                                     (bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & est_black_res_48 > 0 & bk_res_5 > 0 ~ 
                                       "Black Residency Likely Since 1948 & Migration",
                                      (bk_res_1 > 0 | bk_res_2 > 0 | bk_res_3 > 0) & est_black_res_48 == 0 & bk_res_5 > 0 ~ 
                                       "Likely No Black Residency in 1948 & Migration",
                                     TRUE ~ "No Black Residency in 1950")) %>% 
  mutate(zoning = case_when(hazardous > 0  ~ "At Least in Part Deemed Hazardous", 
                            def_decl > 0 ~ "At Least in Part Deemed Definitely Declining & No Part Hazardous", 
                            best > 0 ~ "At Least in Part Deemed Best & No Part Hazardous or Declining", 
                            commercial > 0 | industrial > 0 ~ "Commercial or Industrial", 
                            TRUE ~ "Not Classified"))


#MAKE SURE WE KEEP A LOOK OUT FOR AK 29 - something fishy 
# ggplot(data = df %>% filter(year == 1950 ))+ 
#   geom_sf(aes(fill = ifelse(GRID_ID == "AK-29", 1, 0)))

```

# Plot of Homicide Rate Over Time


```{r, echo=FALSE}

years <- seq(1880, 1965, 5)

year_select <- years[1]

df %>% st_drop_geometry() %>% filter(hom_rt_mean_5>0) %>% 
  summarise(
            q_99_rt = quantile(hom_rt_mean_5, 0.99, na.rm = T), 
            q_999_rt = quantile(hom_rt_mean_5, 0.999, na.rm = T), 
            q_99_ct = quantile(hom_ct_sum_5, 0.99, na.rm = T), 
            q_999_ct = quantile(hom_ct_sum_5, 0.999, na.rm = T), 
            max = max(hom_ct_sum_5, na.rm = T)) 



for (year_select in years) {
  
  g <- ggplot(data = df %>% filter(year == year_select) %>% 
                mutate(hom_rt_mean_5 = ifelse(is.na(hom_rt_mean_5), 0, hom_rt_mean_5))) +
    geom_sf(aes(fill = hom_rt_mean_5, color = as.factor(in_check))) +
    theme_minimal() + 
    labs(title = glue("Homicide Rate {year_select-4}-{year_select}"), 
         fill = "Average Homicide Rate", color = "In Checkerboard") + 
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 3), 
      na.value = "white") 
  
  plot(g)

  
}

for (year_select in years) {
  
  g <- ggplot(data = df %>% filter(year == year_select)) +
    geom_sf(aes(fill = hom_ct_sum_5, color = as.factor(in_check))) +
    theme_minimal() +
    labs(title = glue("Number of Homicides {year_select-4}-{year_select}"), 
         fill = "Total Number of Homicides",  color = "In Checkerboard") +
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 80), 
      na.value = "white") 


  plot(g)

  
}

for (year_select in years) {
  
  g <- ggplot(data = df %>% filter(year == year_select)) +
    geom_sf(aes(fill = bldg_ct_mean_5, color = as.factor(in_check))) +
    theme_minimal() + 
    labs(title = glue("Building Count {year_select-4}-{year_select}"), 
         fill = "Average Building Count", color = "In Checkerboard") + 
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 1700), 
      na.value = "white") 
  
  plot(g)

  
}


 
```

# Zoning as of 1935/40 


```{r, echo=FALSE}

years <- seq(1880, 1965, 5)

year_select <- years[1]


values <- df %>% st_drop_geometry() %>% distinct(zoning) %>% pull()
plot_colors = setNames(c("red", "orange", "green", "black",  "darkgrey"), 
                         c("At Least in Part Deemed Hazardous", 
                         "At Least in Part Deemed Definitely Declining & No Part Hazardous", 
                         "At Least in Part Deemed Best & No Part Hazardous or Declining", 
                         "Commercial or Industrial", 
                         "Not Classified"))



for (year_select in years) {
  
  g <- ggplot(data = df %>% filter(year == year_select) ) +
    geom_sf(aes(fill = hom_rt_mean_5, color = zoning)) +
    theme_minimal() + 
    labs(title = glue("Homicide Rate {year_select-4}-{year_select}"), 
         fill = "Average Homicide Rate", color = "Zoning ca. 1935-40") + 
    scale_color_manual(values = plot_colors) + 
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 3), 
      na.value = "white") 
  
  plot(g)

  
}

for (year_select in years) {
  
  g <- ggplot(data = df %>% filter(year == year_select)) +
    geom_sf(aes(fill = hom_ct_sum_5, color = zoning)) +
    theme_minimal() +
    labs(title = glue("Number of Homicides {year_select-4}-{year_select}"), 
         fill = "Total Number of Homicides",  color = "Zoning ca. 1935-40") +
    scale_color_manual(values = plot_colors) + 
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 80), 
      na.value = "white") 


  plot(g)

  
}



 
```

# Preliminary Identification of Changes 

Calculate the change of total homicides over 5 year periods, the change of average homicides over 5 year periods, and the change of average homicide rate over 5 year periods. 

Define Notable Decline or Rise as change in the upper 10 percent of declines or rises within year. 

```{r, echo=FALSE}

df_changes <- df %>% 
  st_drop_geometry() %>% 
  #remove years with lack of data
  filter(year_5 != 1880 & year_5 != 1930 & year_5 != 1935 & year_5 != 1940 & year != 1945 ) %>% 
  #filter to period of interest 
  filter(year_5 <= 1965) %>%
  distinct(GRID_ID, year_5, hom_ct_sum_5, hom_ct_mean_5, hom_rt_mean_5) %>% 
  arrange(year_5) %>% 
  group_by(GRID_ID) %>% 
  mutate(across(c(hom_ct_sum_5, hom_ct_mean_5, hom_rt_mean_5), ~(. - lag(.)))) %>% 
  filter(year_5 > 1875) %>% 
  ungroup() 



change_calc <- function(var, df = df_changes, decline = TRUE) {

  df["change"] <- df[var]
  
  df <- df %>% 
      distinct(GRID_ID, year_5, change) 
  
  if (decline) {
    df <- df %>%  
      filter(change < 0) %>% 
      arrange(change) %>% 
      group_by(year_5) %>% 
      mutate(pctile = percent_rank(desc(change))) %>%
      mutate(type = glue("decline_{var}")) 
  } else {
    df <- df %>%
      filter(change > 0) %>% 
      arrange(desc(change))%>% 
      group_by(year_5)%>% 
      mutate(pctile = percent_rank(change)) %>%
      mutate(type = glue("rise_{var}")) 
    
  }
  
  df <- df %>% 
    ungroup() %>% 
    filter(pctile > 0.9) 
  
  return(df)
  
}

decline_sum_ct <- change_calc("hom_ct_sum_5", decline = T) 

decline_mean_ct <- change_calc("hom_ct_mean_5", decline = T)

decline_mean_rt <- change_calc( "hom_rt_mean_5", decline = T)
  

rise_sum_ct <- change_calc("hom_ct_sum_5", decline = F) 

rise_mean_ct <- change_calc("hom_ct_mean_5", decline = F)

rise_mean_rt <- change_calc("hom_rt_mean_5", decline = F)


df_changes_sum <- rbind(decline_sum_ct, decline_mean_ct, decline_mean_rt, rise_sum_ct, rise_mean_ct, rise_mean_rt)

```


## Changes Over Time 

```{r, echo=FALSE}

df_geom <- df %>% dplyr::select(GRID_ID, year, in_check) 

df_changes_plot <- inner_join(df_geom, df_changes, by= c("GRID_ID", "year" = "year_5"))

years <- seq(1885, 1965, 5)

year_select <- years[1]

for (year_select in years) {
  
  g <- ggplot(data = df_changes_plot %>% filter(year == year_select) ) +
    geom_sf(aes(fill = hom_rt_mean_5, color = as.factor(in_check))) +
    theme_minimal() + 
    labs(title = glue("Change in Homicide Rate Between {year_select-9}-{year_select-5} and {year_select-4}-{year_select}"), 
         fill = "Change in Average Homicide Rate", color = "In Checkerboard") + 
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 3), 
      na.value = "white") 
  
  plot(g)

  
}

for (year_select in years) {
  
  g <- ggplot(data = df_changes_plot %>% filter(year == year_select)) +
    geom_sf(aes(fill = hom_ct_sum_5, color = as.factor(in_check))) +
    theme_minimal() +
    labs(title = glue("Change in Total Number of Homicides Between {year_select-9}-{year_select-5} and {year_select-4}-{year_select}"), 
         fill = "Change in Total Number of Homicides",  color = "In Checkerboard") +
    scale_color_manual(values = c("darkgrey", "black")) +
    scale_fill_gradientn(
      colors = c("white",  "#7ce8ff", "#55d0ff", "#00acdf","#0080bf","#00498D", "#02386E", "#00264D"), 
      limits = c(0, 80), 
      na.value = "white") 


  plot(g)


}

 
```

## Notable Changes (Upper 10% of Rises or Declines in a Given Year)
### 1950s - 1960s

```{r, echo=FALSE}

year_c <- 1955
var <- "hom_ct_sum_5"

plot_rise <- function(year_c, var) {
  
  title_piece <- ifelse(var == "hom_ct_sum_5", "Total Number of Homicides", "Average Homicide Rate")

  #Get riots in that time period
  riots <- df %>% 
    st_drop_geometry() %>% 
    filter(!is.na(name_riot)) %>% distinct(GRID_ID, year) %>% 
    filter(year >= year_c-5 & year <= year_c) %>% 
    pull(GRID_ID)
  
  #Get significant increases in that time period
  IDS_rise <- df_changes_sum %>% 
    filter(type == glue("rise_{var}"),year_5 == year_c) %>% 
    distinct(GRID_ID) %>% 
    pull()
  
  IDS_decline <- df_changes_sum %>% 
    filter(type == glue("decline_{var}"),year_5 == year_c) %>% 
    distinct(GRID_ID) %>% 
    pull()
  
  
  #Filter data to those IDS
  
  df_plot <- df %>% 
    filter(year == year_c) %>%
    mutate(flag_rise = ifelse(GRID_ID %in% IDS_rise,  "Notable Rise", NA), 
           flag_decline = ifelse(GRID_ID %in% IDS_decline , "Notable Decline", NA), 
         flag_riot = ifelse(GRID_ID %in% riots , "Riot", NA)) %>% 
    mutate(flag = paste0(flag_rise, ifelse(is.na(flag_rise), "", ifelse(!is.na(flag_decline) | 
                                                                          !is.na(flag_riot) | 
                                                                          !is.na(bk_res_since_39)," & ", "")),
                        flag_decline, ifelse(is.na(flag_decline), "", ifelse(!is.na(flag_riot) | 
                                                                          !is.na(bk_res_since_39)," & ", "")), 
                         flag_riot, ifelse(is.na(flag_riot), "", ifelse(!is.na(bk_res_since_39), " & ", "")), 
                        bk_res_since_39), 
           flag = str_replace_all(flag, "NA", "")) %>% 
    filter(!is.na(flag)) %>% 
    mutate(bk_res_4_cat = ifelse(is.na(bk_res_4_cat), "No Change", bk_res_4_cat))
  
  
  df_plot %>% st_drop_geometry() %>% distinct(flag)
  
    #Labels
  
   plot_colors <- setNames(c( #"Notable Rise"
                              "#c39797", 
                              #Notable Rise & Black Residency Likely Since 1939 & Migration
                             "#632523",
                             #Notable Rise & Likely No Black Residency in 1939 & Migration
                             "#2a0000",
                             
                             #Notable Rise & Riot
                             "red",
                             #Notable Rise & Black Residency Likely Since 1939 & No Migration
                             "#ffc04c", 
                             #Notable Rise & No Black Residency in 1950
                             "#ffdfbf",
                             
                             
                             #Notable Decline
                            "#A8DCAB", 
                            #Notable Decline & Black Residency Likely Since 1939 & Migration
                            "#80EF80", 
                            #Notable Decline & Likely No Black Residency in 1939 & Migration
                            "#05472A",
                            
                            
                            #Notable Decline & Black Residency Likely Since 1939 & No Migration
                            "green",
                            #Notable Decline & No Black Residency in 1950 
                            "darkgreen",
                            
                            #Black Residency Likely Since 1939 & Migration
                            "#6897bb", 
                            #Likely No Black Residency in 1939 & Migration
                            "lightblue", 
                            #Black Residency Likely Since 1939 & No Migration
                            "#254061", 
                            #No Black Residency in 1950
                            "#D3D3D3", 
                            
                            #Riot 
                            "maroon", 
                            #Notable Decline & Riot
                            "#f23fb2", 
                            #Riot & No Black Residency in 1950
                            "purple", 
                            #"Riot & Black Residency Likely Since 1939 & No Migration",
                            "black",
                            #"Riot & Black Residency Likely Since 1939 & Migration",
                            "red",
                            #"Riot & Likely No Black Residency in 1939 & Migration",
                            "#e66fa0",
                            
                            #Notable Rise & Riot & Black Residency Likely Since 1939 & Migration
                            "black", 
                            #Notable Rise & Riot & Black Residency Likely Since 1939 & No Migration
                            "black",  
                            #Notable Rise & Riot & Likely No Black Residency in 1939 & Migration
                            "black",
                            #Notable Rise & Riot & No Black Residency in 1950
                            "black", 
                            
                            #Notable Decline & Riot & Black Residency Likely Since 1939 & Migration
                            "black", 
                            #Notable Decline & Riot & Black Residency Likely Since 1939 & No Migration
                            "black",  
                            #Notable Decline & Riot & Likely No Black Residency in 1939 & Migration
                            "black",
                            #Notable Decline & Riot & No Black Residency in 1950
                            "black"
                            ),
                          c("Notable Rise", 
                            "Notable Rise & Black Residency Likely Since 1939 & Migration", 
                            "Notable Rise & Likely No Black Residency in 1939 & Migration", 
                            
                            "Notable Rise & Riot",
                           "Notable Rise & Black Residency Likely Since 1939 & No Migration", 
                           "Notable Rise & No Black Residency in 1950", 
                          
                          "Notable Decline", 
                          "Notable Decline & Black Residency Likely Since 1939 & Migration",  
                          "Notable Decline & Likely No Black Residency in 1939 & Migration",  
                          
                          "Notable Decline & Black Residency Likely Since 1939 & No Migration", 
                          "Notable Decline & No Black Residency in 1950", 
                          
                          "Black Residency Likely Since 1939 & Migration", 
                          "Likely No Black Residency in 1939 & Migration",
                          "Black Residency Likely Since 1939 & No Migration", 
                          "No Black Residency in 1950", 
                          
                          "Riot",  
                          "Notable Decline & Riot",
                          "Riot & No Black Residency in 1950", 
                          "Riot & Black Residency Likely Since 1939 & No Migration",
                          "Riot & Black Residency Likely Since 1939 & Migration",
                          "Riot & Likely No Black Residency in 1939 & Migration",
                          
                          "Notable Rise & Riot & Black Residency Likely Since 1939 & Migration",  
                          "Notable Rise & Riot & Black Residency Likely Since 1939 & No Migration", 
                          "Notable Rise & Riot & Likely No Black Residency in 1939 & Migration", 
                          "Notable Rise & Riot & No Black Residency in 1950",
                          
                          "Notable Decline & Riot & Black Residency Likely Since 1939 & Migration",  
                          "Notable Decline & Riot & Black Residency Likely Since 1939 & No Migration", 
                          "Notable Decline & Riot & Likely No Black Residency in 1939 & Migration", 
                          "Notable Decline & Riot & No Black Residency in 1950")) 
   
  
  plot_colors <- plot_colors[df_plot %>% distinct(flag) %>% pull()]
  
  plot_colors_lines <- setNames(c( "#e5ff4c", "#ff7f00","#836FFF",
                                  "white"),
                                c("Any Increase From 1% or More to More than 90%",
                                  "Significant Increase From 1% or More to Less than 90%",
                                  "Any Increase From Less Than 1%", "No Change"))


  plot_colors_lines <- plot_colors_lines[df_plot %>% distinct(bk_res_4_cat)  %>% pull()]


  g <- ggplot(data = df_plot) +
      geom_sf(aes(fill = flag)) +
      theme_minimal() + 
      labs(title =  glue("Change in {title_piece} Between {year_c - 9}-{year_c - 5} and {year_c - 4}-{year_c}"), 
           subtitle = glue("Intersection with Black Residency as of 1939 and 1950 & Riot in {year_c - 4}-{year_c}")) + 
      scale_fill_manual(values = plot_colors) 
    
  plot(g)
  
  # #Plot
  # g <- ggplot(data = df_plot) +
  #     geom_sf(aes(fill = flag, color = bk_res_4_cat), linewidth = 0.55) +
  #     theme_minimal() + 
  #     labs(title =  glue("Change for {year_c}"))+ 
  #     scale_fill_manual(values = plot_colors) +
  #     scale_color_manual(values = plot_colors_lines) 
  #   
  # plot(g)
  
  df_plot %>% 
    st_drop_geometry() %>% 
    group_by(flag) %>% 
    summarise(count = n()) %>% 
    ungroup() %>% 
    arrange(flag)
  
}



plot_rise(1950, "hom_ct_sum_5")

plot_rise(1955, "hom_ct_sum_5")

plot_rise(1960, "hom_ct_sum_5")

plot_rise(1965, "hom_ct_sum_5")



plot_rise(1950, "hom_rt_mean_5")

plot_rise(1955, "hom_rt_mean_5")

plot_rise(1960, "hom_rt_mean_5")

plot_rise(1965, "hom_rt_mean_5")

```



# Investigate Change in Black Residency 1939 - 1848 - 1950 and Zoning 



## Investigate Riot Years 



# Correlations 


## Rough Correlation Check 

```{r, echo = FALSE}
df_no_geom <- df %>% st_drop_geometry()

calc_cor <- function(var, control_var = "hom_ct_sum_5", data = df_no_geom) {
  print(var)
 correlation <-  cor(data[control_var], data[var], use = "complete.obs")[1]
 output <- data.frame(variable = var, 
                      correlation = correlation)
 return(output)
}

cols <- c("year_riot", "in_check", "still_desi",  "def_decl", "hazardous" , "industrial", "best", "commercial" , 
          "bk_res_1" , "bk_res_2", "bk_res_3", "bk_res_4_1", "bk_res_4_2" , "bk_res_4_3" , "bk_res_5_1" ,  "bk_res_5_2", 
          "bk_res_5_3" )
output <- map_df(cols, calc_cor)

output %>% 
  arrange(-abs(correlation))

```

# Testing 


## Testing Black Residency Values


We see polygons where the percentage of 1% concentration is lower than 10% or 90%. There are very few instances where the difference is large (more than 10 percentage points difference in area mapped), and most occur in the south of Chicago where the maps were a) most distorted and b) the least amount of development matches the present state. 

It may be best to use the 90% value rather than the 10% or 1% value in these instances. This is further corroborated by the original maps. The areas plotted below where the difference is highest appear to be of fairly equal size in all three maps. 


```{r}


test <- df %>%
  filter(year == 1950) %>% 
  filter(bk_res_3 < bk_res_2 | bk_res_2 < bk_res_1 | bk_res_3 < bk_res_1) %>%
  dplyr::select(GRID_ID, bk_res_1, bk_res_2, bk_res_3) %>% 
  mutate(diff_1_2 = bk_res_1 - bk_res_2, 
         diff_1_3 = bk_res_1 - bk_res_3, 
         diff_2_3 = bk_res_2 - bk_res_3) %>% 
  pivot_longer(cols = contains("diff")) %>% 
  filter(value > 0.1) 

test 

ggplot(data = test ) + 
  geom_sf(aes(fill = GRID_ID)) + 
  coord_sf(
    xlim = c(-87.94011, -87.52398), # Approximate longitude bounds of Chicago
    ylim = c(41.64454, 42.02304)    # Approximate latitude bounds of Chicago
  ) 
  
```
