---
title: "01_homicide_change_and_black_migr"
output: html_document
---

```{r, include = FALSE}
library(gganimate)
library(magick)
library(fillr)
library(ggrepel)
library(av)
library(zoo)

rm(list = ls())
source("header.R")

#Hexagons 
hex <- st_read("data/intermediate/homicide_hexagons/survivorship")
df <- st_read("data/mst/hex_merged_w_black_migr.gpkg") 

#Streets
streets <- st_read("data/intermediate/highways.gpkg")

#Projects and Riots

cha_projects <- read_sf("data/raw/housing_projects/housing_v3") 

riots <- read_sf("data/intermediate/racial_violence/riots.geojson") %>% 
  mutate(name_clean = str_remove(name, " housing project") ,
         label = paste0("Rt:", year))
cha_projects <- st_transform(cha_projects, st_crs(hex))
riots <- st_transform(riots, st_crs(hex))




```

# CHANGES 1939 MIGRATION --------------------------------

## Year-by-Year Time Series -----------------------------------

```{r}


df_changes <- df %>% 
  st_drop_geometry() %>% 
  distinct(GRID_ID, year, hom_ct, hom_rt, name_prj, bldst_prj, name_riot, year_riot, bk_res_since_39, in_check) %>% 
  arrange(year) %>% 
  group_by(GRID_ID) %>% 
  mutate(across(c(hom_ct), ~((. - lag(.))/lag(.)))) %>% 
  ungroup() 

year_start <- 1940 
year_end <- 1955

df_main_lim <- df %>% 
  dplyr::select(GRID_ID, year) %>%
  left_join(df_changes, by = c("GRID_ID", "year")) %>% 
  filter(str_detect(bk_res_since_39, "& Black Migration")) %>% 
  filter(year >= year_start , year <= year_end)

riots_expand <- riots %>% 
  st_drop_geometry() %>% 
  group_by(name) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>% 
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(id)) %>% 
  ungroup()

riots_expand <- riots %>% 
  dplyr::select(name) %>% 
  left_join(riots_expand, by = "name") %>% 
  filter(year <= year_end) %>% 
  mutate(year = as.integer(year))

housing_expand <- cha_projects %>% 
  st_drop_geometry() %>% 
  mutate(year = opened) %>% 
  group_by(name) %>%
  complete(year = seq(opened, closed, 1)) %>% 
  arrange(year) %>%
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(bldst)) %>% 
  ungroup()

housing_expand <- cha_projects %>% 
  dplyr::select(name) %>% 
  left_join(housing_expand, by = "name") %>% 
  filter(year <= year_end & year >= year_start) %>% 
  mutate(year = as.integer(year))

streets_select <- streets %>% 
  st_drop_geometry() %>% 
  filter(DEDICATED_ %in% c("EISENHOWER EXPY", 
                           "ADLAI STEVENSON EXPY", 
                           "CHICAGO SKWY", 
                            "DAN RYAN EXPY", 
                           "KENNEDY EXPY")) %>% 
  mutate(year = begin) %>% 
  group_by(OBJECTID) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>%
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(begin)) %>% 
  ungroup() %>% 
  dplyr::distinct(OBJECTID, DEDICATED_, year, begin, open)

streets_select <- streets %>% 
  dplyr::select(OBJECTID) %>% 
  left_join(streets_select, by = "OBJECTID") %>% 
  filter(year <= year_end & year >= year_start) %>% 
  mutate(year = as.integer(year))

```


```{r}



label_positions_riot <- st_centroid(riots_expand) %>%
  st_transform(4326) %>%  # Convert back to WGS84 if needed
  cbind(st_coordinates(.))

label_positions_housing <-  st_centroid(housing_expand) %>%
  st_transform(4326) %>%  # Convert back to WGS84 if needed
  cbind(st_coordinates(.))

g<- ggplot() + 
  transition_time(year) + 
  geom_sf(data = df_main_lim , aes(fill = hom_ct, color = as.factor(in_check))) + 
  geom_sf(data = riots_expand, size = 1, color = "green") + 
  geom_sf(data = housing_expand, color = "purple", size= 0.3) +
  geom_sf(data = streets_select, aes(color =  paste0(DEDICATED_,": ", begin, "-", open))) + 
  geom_text(data = label_positions_riot, position=position_jitter(), 
            aes(x = X, y = Y, label = label, vjust = 1), size = 1) +
  geom_text_repel(data = label_positions_housing, position=position_jitter(), 
            aes(x = X, y = Y, label = label, vjust = 1), size = 1) +
  labs(title = glue("Hexagons with Black Migration from 1939 to 1950 & Change in Homicides"), 
       subtitle ="Year: {frame_time}",
         fill = "Change of Homicides",  color = "Highways & In Checkerboard", x = "", y = "") +
  scale_fill_gradientn(
      colors = c( "#00264D", "#55d0ff", "#7ce8ff","white", "pink", "red", "darkred"  ), 
      limits = c(-1.5, 1.5), 
      na.value = "white") +
  scale_color_manual(values = c("darkgrey", "black", "pink")) + 
  theme_minimal() +
  coord_sf(
    xlim = c(-87.746, -87.5), 
    ylim = c(41.6, 42.0)    
  ) +
  theme(legend.key.size = unit(0.4, 'cm')) +
  theme(legend.title = element_text(size = 4), 
        legend.text  = element_text(size = 4),  
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        plot.title = element_text(size = 8, face = "bold"),
        plot.subtitle = element_text(size = 6))

anim <- gganimate::animate(g, renderer = av_renderer("output/year_by_year_change_39_perc.mp4"), 
                fps = 0.5, nframes = 16, 
                width = 1920,
                height = 1080,
                res = 300, 
                dir = "ouput/")


#magick::image_write(anim, path="myanimation.gif")
  
```


How do we account for high homicide places. Higher homicide level --> more fluctuation. 

Look into when which segment of Eisenhower was built - there appear to be some segments that see immediate increase and some that see decline 

think about how to de-mean data: maybe remove year average for those neighborhoods? for entire city? 

Create a chart showing 5 year-out average values for each year 

Create a measure that looks at both year over year variance - is there value in knowing that some hexagons fluctuate more over time than others? 

What makes something a top increase or decrease? How do we account for high homicide and low homicide areas? 

How do we account for fluctuation - fluctuation is an inherent property but assumably waves make a decrease for high homicide areas quite likely 

Thoughts regression: 

- Tri-hexagons? choose those hexagons that are on border to which housing is closest to be treated 
- Assumably already working with clusters of common trends 
- Concerned about variation of high homicide areas even accounting for a) entity fixed effects b) year fixed effects c) proportional measures 


## Year-by-Year 5-Year Out Average ------------------------------

```{r}


df_changes <- df %>% 
   st_drop_geometry() %>% 
  distinct(GRID_ID, year, hom_ct, hom_rt, name_prj, bldst_prj, name_riot, year_riot, bk_res_since_39, in_check) %>% 
  group_by(GRID_ID) %>%
  arrange(year) %>%
  mutate(
        # 5-year forward-looking average (current year + next 4)
        rolling_avg = rollapply(hom_ct, 
                              width = 5, 
                              FUN = \(x) mean(x, na.rm = TRUE),
                              align = "left",
                              fill = NA),
        
        # Yearly change in rolling average
        yoy_change = rolling_avg - lag(rolling_avg)
    ) %>%
    ungroup()

year_start <- 1940 
year_end <- 1955

df_main_lim <- df %>% 
  dplyr::select(GRID_ID, year) %>%
  left_join(df_changes, by = c("GRID_ID", "year")) %>% 
  filter(str_detect(bk_res_since_39, "& Black Migration")) %>% 
  filter(year >= year_start , year <= year_end)


riots_expand <- riots %>% 
  st_drop_geometry() %>% 
  group_by(name) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>% 
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(id)) %>% 
  ungroup()

riots_expand <- riots %>% 
  dplyr::select(name) %>% 
  left_join(riots_expand, by = "name") %>% 
  filter(year <= year_end) %>% 
  mutate(year = as.integer(year))

housing_expand <- cha_projects %>% 
  st_drop_geometry() %>% 
  mutate(year = bldst) %>% 
  group_by(name) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>%
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(bldst)) %>% 
  ungroup()

housing_expand <- cha_projects %>% 
  dplyr::select(name) %>% 
  left_join(housing_expand, by = "name") %>% 
  filter(year <= year_end & year >= year_start) %>% 
  mutate(year = as.integer(year))

streets_select <- streets %>% 
  st_drop_geometry() %>% 
  filter(DEDICATED_ %in% c("EISENHOWER EXPY", 
                           "ADLAI STEVENSON EXPY", 
                           "CHICAGO SKWY", 
                            "DAN RYAN EXPY", 
                           "KENNEDY EXPY")) %>% 
  mutate(year = begin) %>% 
  group_by(OBJECTID) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>%
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(begin)) %>% 
  ungroup() %>% 
  dplyr::distinct(OBJECTID, DEDICATED_, year, begin, open)

streets_select <- streets %>% 
  dplyr::select(OBJECTID) %>% 
  left_join(streets_select, by = "OBJECTID") %>% 
  filter(year <= year_end & year >= year_start) %>% 
  mutate(year = as.integer(year))

```


```{r}


label_positions_riot <- st_centroid(riots_expand) %>%
  st_transform(4326) %>%  # Convert back to WGS84 if needed
  cbind(st_coordinates(.))

label_positions_housing <-  st_centroid(housing_expand) %>%
  st_transform(4326) %>%  # Convert back to WGS84 if needed
  cbind(st_coordinates(.))

g<- ggplot() + 
  transition_time(year) + 
  geom_sf(data = df_main_lim , aes(fill = yoy_change)) + 
  geom_sf(data = riots_expand, size = 1, color = "green") + 
  geom_sf(data = housing_expand, color = "purple", size= 0.3) +
  geom_sf(data = streets_select, aes(color =  paste0(DEDICATED_,": ", begin, "-", open))) + 
  geom_text(data = label_positions_riot, position=position_jitter(), 
            aes(x = X, y = Y, label = label, vjust = 1), size = 1) +
  geom_text_repel(data = label_positions_housing, position=position_jitter(), 
            aes(x = X, y = Y, label = label, vjust = 1), size = 1) +
  labs(title = glue("Hexagons with Black Migration from 1948 to 1950 & Change of Homicides"), 
       subtitle ="Year: {frame_time}",
         fill = "Change of Homicides",  color = "Highways", x = "", y = "") +
  scale_fill_gradientn(
      colors = c( "#00264D", "#55d0ff", "#7ce8ff","white", "pink", "red", "darkred"  ), 
      limits = c(-1.5, 1.5), 
      na.value = "white") +
  theme_minimal() +
  coord_sf(
    xlim = c(-87.746, -87.5), 
    ylim = c(41.6, 42.0)    
  ) +
  theme(legend.key.size = unit(0.4, 'cm')) +
  theme(legend.title = element_text(size = 4), 
        legend.text  = element_text(size = 4),  
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        plot.title = element_text(size = 8, face = "bold"),
        plot.subtitle = element_text(size = 6))

anim <- gganimate::animate(g, renderer = av_renderer("output/5year_by_5year_change_39.mp4"), 
                fps = 0.5, nframes = 16, 
                width = 1920,
                height = 1080,
                res = 300)

```
# CHANGES 1948 MIGRATION --------------------------------

## Year-by-Year Time Series -----------------------------------

```{r}


df_changes <- df %>% 
  st_drop_geometry() %>% 
  distinct(GRID_ID, year, hom_ct, hom_rt, name_prj, name_riot, year_riot, bk_res_since_48) %>% 
  arrange(year) %>% 
  group_by(GRID_ID) %>% 
  mutate(across(c(hom_ct), ~((. - lag(.))/lag(.)))) %>% 
  ungroup() 

year_start <- 1948
year_end <- 1955

df_main_lim <- df %>% 
  dplyr::select(GRID_ID, year) %>%
  left_join(df_changes, by = c("GRID_ID", "year")) %>% 
  filter(str_detect(bk_res_since_48, "& Black Migration")) %>% 
  filter(year >= year_start , year <= year_end)

riots_expand <- riots %>% 
  st_drop_geometry() %>% 
  group_by(name) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>% 
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(id)) %>% 
  ungroup()

riots_expand <- riots %>% 
  dplyr::select(name) %>% 
  left_join(riots_expand, by = "name") %>% 
  filter(year <= year_end) %>% 
  mutate(year = as.integer(year))

housing_expand <- cha_projects %>% 
  st_drop_geometry() %>% 
  mutate(year = bldst) %>% 
  group_by(name) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>%
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(bldst)) %>% 
  ungroup()

housing_expand <- cha_projects %>% 
  dplyr::select(name) %>% 
  left_join(housing_expand, by = "name") %>% 
  filter(year <= year_end & year >= year_start) %>% 
  mutate(year = as.integer(year))

streets_select <- streets %>% 
  st_drop_geometry() %>% 
  filter(DEDICATED_ %in% c("EISENHOWER EXPY", 
                           "ADLAI STEVENSON EXPY", 
                           "CHICAGO SKWY", 
                            "DAN RYAN EXPY", 
                           "KENNEDY EXPY")) %>% 
  mutate(year = begin) %>% 
  group_by(OBJECTID) %>%
  complete(year = seq(year_start, year_end, 1)) %>% 
  arrange(year) %>%
  fill(everything(), .direction = "down") %>% 
  filter(!is.na(begin)) %>% 
  ungroup() %>% 
  dplyr::distinct(OBJECTID, DEDICATED_, year, begin, open)

streets_select <- streets %>% 
  dplyr::select(OBJECTID) %>% 
  left_join(streets_select, by = "OBJECTID") %>% 
  filter(year <= year_end & year >= year_start) %>% 
  mutate(year = as.integer(year))

```


```{r}



label_positions_riot <- st_centroid(riots_expand) %>%
  st_transform(4326) %>%  # Convert back to WGS84 if needed
  cbind(st_coordinates(.))

label_positions_housing <-  st_centroid(housing_expand) %>%
  st_transform(4326) %>%  # Convert back to WGS84 if needed
  cbind(st_coordinates(.))

g<- ggplot() + 
  transition_time(year) + 
  geom_sf(data = df_main_lim , aes(fill = hom_ct)) + 
  geom_sf(data = riots_expand, size = 1, color = "green") + 
  geom_sf(data = housing_expand, color = "purple", size= 0.3) +
  geom_sf(data = streets_select, aes(color =  paste0(DEDICATED_,": ", begin, "-", open))) + 
  geom_text(data = label_positions_riot, position=position_jitter(), 
            aes(x = X, y = Y, label = label, vjust = 1), size = 1) +
  geom_text_repel(data = label_positions_housing, position=position_jitter(), 
            aes(x = X, y = Y, label = label, vjust = 1), size = 1) +
  labs(title = glue("Hexagons with Black Migration from 1948 to 1950 & Change of Homicides"), 
       subtitle ="Year: {frame_time}",
         fill = "Change of Homicides",  color = "Highways", x = "", y = "") +
  scale_fill_gradientn(
      colors = c( "#00264D", "#55d0ff", "#7ce8ff","white", "pink", "red", "darkred"  ), 
      limits = c(-1.5, 1.5), 
      na.value = "white") +
  theme_minimal() +
  coord_sf(
    xlim = c(-87.746, -87.5), 
    ylim = c(41.6, 42.0)    
  ) +
  theme(legend.key.size = unit(0.4, 'cm')) +
  theme(legend.title = element_text(size = 4), 
        legend.text  = element_text(size = 4),  
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        plot.title = element_text(size = 8, face = "bold"),
        plot.subtitle = element_text(size = 6))

anim <- gganimate::animate(g, renderer = av_renderer("output/year_by_year_change_48_perc.mp4"), 
                fps = 0.5, nframes = 16, 
                width = 1920,
                height = 1080,
                res = 300)


#magick::image_write(anim, path="myanimation.gif")
  
```






