---
title: "02_west_side_patterns"
output: html_document
---



```{r Set Up, include = FALSE}
rm(list = ls())
source("header.R")
library(geosphere)
library(lubridate)
library(tidygeocoder)
library(mapview)
library(mapedit)

# HOMICIDES 

  ## Pre 65 
homs <- st_read("data/raw/homicides_1940-1965_geocoded_v3") %>% 
  mutate(year = as.integer(year)) %>%
  mutate(category_clean = case_when(str_detect(tolower(cod), "police") ~ "Police Killing",
                                    str_detect(tolower(cod), "gun") |  str_detect(tolower(cod), "shot") 
                                    |  str_detect(tolower(cod), "bullet") ~ "Gunshot", 
                                    str_detect(tolower(cod), "stab") | str_detect(tolower(cod), "cut") ~ "Knife", 
                                   TRUE ~ "Other"), 
         date_clean = as_date(date), 
         category = paste0(tolower(cod), ": ", format(date_clean, "%m/%d"))) %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!str_detect(category, "burn"))%>% 
  filter(!str_detect(category, "overdose"))%>% 
  filter(!str_detect(category, "abortion"))%>% 
  filter(!str_detect(category, "fire"))

homs_clean <- homs %>% 
  dplyr::select(year, category)

grid <- st_read("data/raw/Fishnet_InsideChicago_150m")

post_65 <- read_csv("data/raw/IndividualHomicides_1965_2022.csv") %>%
  filter(!is.na(X_MetCent) & !is.na(Y_MetCent))
post_65 <- st_as_sf(post_65, coords = c("X_MetCent", "Y_MetCent"), crs = st_crs(grid))
post_65 <- st_transform(post_65, crs = st_crs(homs))

post_65_clean <- post_65 %>%
  dplyr::select(year) %>%
  mutate(category= "")

census_40_tract <- st_read("data/raw/census/1940_census_tracts")
census_40_tract <- st_transform(census_40_tract, crs = st_crs(homs))
  ##combine 

homs_all <- rbind(homs_clean, post_65_clean)
#homs_all <- homs_clean


#Other data 

cha_projects <- read_sf("data/intermediate/housing_projects/cha_projects.geojson") %>% 
  rename(bldst = yr_built_start, bldend = yr_built_end, dmlst = yr_dmlsh_or_cnvrt_start, dmlend = yr_dmlsh_or_cnvrt_end ) %>% 
  mutate(label = paste0("HP:", bldst-1900, ifelse(!is.na(bldend), paste0("-", bldend-1900), "")))
cha_projects <- st_transform(cha_projects, st_crs(homs))
 ##Post 65

riots <- read_sf("data/intermediate/racial_violence/riots.geojson") %>% 
  mutate(name_clean = str_remove(name, " housing project") ,
         label = paste0("Rt:", year))

riots <- st_transform(riots, st_crs(homs))



#Neighborhoods

westside <- read_sf("data/intermediate/westside.geojson")
westside <- st_transform(westside, st_crs(homs))

nw1 <- read_sf("data/mst/neighborhoods/nw1.shp") %>% 
  mutate(size = st_area(geometry)) %>% 
  mutate(neighborhood = "nw1")
nw2 <- read_sf("data/mst/neighborhoods/nw2.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "nw2")
nc1 <- read_sf("data/mst/neighborhoods/nc1.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "nc1")
ne1 <- read_sf("data/mst/neighborhoods/ne1.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "ne1")
ne2 <- read_sf("data/mst/neighborhoods/ne2.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "ne2")
ne3 <- read_sf("data/mst/neighborhoods/ne3.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "ne3")
e1 <- read_sf("data/mst/neighborhoods/e1.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "e1")
sw1 <- read_sf("data/mst/neighborhoods/sw1.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "sw1")
sc1 <- read_sf("data/mst/neighborhoods/sc1.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "sc1")
sc2 <- read_sf("data/mst/neighborhoods/sc2.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "sc2")
sc3 <- read_sf("data/mst/neighborhoods/sc3.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "sc3")
se1 <- read_sf("data/mst/neighborhoods/se1.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "se1")
se2 <- read_sf("data/mst/neighborhoods/se2.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "se2")
se3 <- read_sf("data/mst/neighborhoods/se3.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "se3")
se4 <- read_sf("data/mst/neighborhoods/se4.shp")%>% 
  mutate(size = st_area(geometry))%>% 
  mutate(neighborhood = "se4")

all_neighborhoods <- nw1 %>% 
  union_all(nw2)%>% 
  union_all(nc1)%>% 
  union_all(ne1)%>% 
  union_all(ne2)%>% 
  union_all(ne3)%>% 
  union_all(e1)%>% 
  union_all(sw1)%>% 
  union_all(sc1)%>% 
  union_all(sc2)%>% 
  union_all(sc3)%>% 
  union_all(se1)%>% 
  union_all(se2)%>% 
  union_all(se3)%>% 
  union_all(se4)

# harrison_cermak <- read_sf("data/mst/neighborhoods/harrison_cermak.shp")
# lower_south_west <- read_sf("data/mst/neighborhoods/lower_south_west.shp")
# north_middle_eisenhower <- read_sf("data/mst/neighborhoods/north_middle_eis.shp")
# upper_north_west <- read_sf("data/mst/neighborhoods/nortwest.shp")
# upper_north_east <- read_sf("data/mst/neighborhoods/upper_north_east.shp")
# south_west_eisenhower <- read_sf("data/mst/neighborhoods/southwest_roos.shp")


```

```{r}

homs_west <- homs_all %>% 
  st_join(westside, join = st_intersects, left = FALSE) %>%
  arrange(year, category) %>% 
  mutate(id = row_number())

# census_clean  <- st_make_valid(census_40_tract)
# westside_clean <- st_make_valid(westside)
# 
# census_tract_40_west <- st_filter(census_clean, westside_clean)
  
```
```{r}
sw1_homs <- homs_west %>% st_join(sw1, join = st_intersects, left = FALSE) %>% pull(id)
sc1_homs <- homs_west %>% st_join(sc1, join = st_intersects, left = FALSE) %>% pull(id)
sc2_homs <- homs_west %>% st_join(sc2, join = st_intersects, left = FALSE) %>% pull(id)
sc3_homs <- homs_west %>% st_join(sc3, join = st_intersects, left = FALSE) %>% pull(id)
se1_homs <- homs_west %>% st_join(se1, join = st_intersects, left = FALSE) %>% pull(id)
se2_homs <- homs_west %>% st_join(se2, join = st_intersects, left = FALSE) %>% pull(id)
se3_homs <- homs_west %>% st_join(se3, join = st_intersects, left = FALSE) %>% pull(id)
se4_homs <- homs_west %>% st_join(se4, join = st_intersects, left = FALSE) %>% pull(id)
e1_homs <- homs_west %>% st_join(e1, join = st_intersects, left = FALSE) %>% pull(id)
nc1_homs <- homs_west %>% st_join(nc1, join = st_intersects, left = FALSE) %>% pull(id)
ne1_homs <- homs_west %>% st_join(ne1, join = st_intersects, left = FALSE) %>% pull(id)
ne2_homs <- homs_west %>% st_join(ne2, join = st_intersects, left = FALSE) %>% pull(id)
ne3_homs <- homs_west %>% st_join(ne3, join = st_intersects, left = FALSE) %>% pull(id)
nw1_homs <- homs_west %>% st_join(nw1, join = st_intersects, left = FALSE) %>% pull(id)
nw2_homs <- homs_west %>% st_join(nw2, join = st_intersects, left = FALSE) %>% pull(id)

homs_west <- homs_west %>% 
  mutate(neighborhood = case_when(id %in% sw1_homs ~ "sw1", 
                                  id %in% sc1_homs ~ "sc1", 
                                  id %in% sc2_homs ~ "sc2", 
                                  id %in% sc3_homs ~ "sc3", 
                                  id %in% se1_homs ~ "se1", 
                                  id %in% se2_homs ~ "se2", 
                                  id %in% se3_homs ~ "se3",
                                  id %in% se4_homs ~ "se4", 
                                  id %in% e1_homs ~ "e1", 
                                  id %in% nc1_homs ~ "nc1", 
                                  id %in% ne1_homs ~ "ne1", 
                                  id %in% ne2_homs ~ "ne2",
                                  id %in% ne3_homs ~ "ne3", 
                                  id %in% nw1_homs ~ "nw1", 
                                  id %in% nw2_homs ~ "nw2", 
                                  TRUE ~ "not categorized"))

homs_neighborhood_count <- homs_west %>% 
  st_drop_geometry() %>%
  group_by(year, neighborhood) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  filter(neighborhood != "not categorized") %>% 
  left_join(all_neighborhoods %>% st_drop_geometry())
```
```{r}

# all_neighborhoods <- nw1 %>%
# st_union(nw2) %>%
# st_union(nc1)%>%
# st_union(ne1)%>%
# st_union(ne2)%>%
# st_union(ne3)%>%
# st_union(e1)%>%
# st_union(sw1)%>%
# st_union(sc1)%>%
# st_union(sc2)%>%
# st_union(sc3)%>%
# st_union(se1)%>%
# st_union(se2)%>%
# st_union(se3)%>%
# st_union(se4)
# 
# leaflet() %>%
#   addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
#   addPolygons(data = all_neighborhoods,  opacity = 0)
# 
# test <- homs_west %>% 
#   st_join(all_neighborhoods, join = st_intersects, left = TRUE) %>% 
#   filter(is.na(ftr_typ))
# 
# 
# leaflet() %>%
#   addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
#   addPolygons(data = all_neighborhoods,  opacity = 0) %>% 
#   addCircleMarkers(
#     data        = test,
#     radius      = 1,
#     color       = "black",
#     fillOpacity = 0.6,
#     label       = ~as.character(year),
#     labelOptions = labelOptions(
#       direction = "auto",
#       textsize  = "14px",
#       opacity   = 0.9,
#       offset    = c(4, -4)
#     ))


```


#INDIVIDUAL MAPS

```{r}


# Set up the years to iterate through

year <- 1955 
center_westside  <- st_coordinates(st_centroid(westside))


map_year <- function(year, data = homs_west, center = center_westside) {
  
  homs_year <- data %>% filter(year == !!year)
  
  nhoods <- unique(homs_year$neighborhood)

  # Calculate standard deviation ellipse for the current year
  sde_data <- std_dev_ellipse(homs_year)
  sde <- st_ellipse(geometry = sde_data, sx = sde_data$sx, sy = sde_data$sy, rotation = -sde_data$theta) 

  sde_sf <- st_sf(geometry = sde)%>% 
    mutate(year = !!year)


# Create leaflet map
  map <- leaflet() %>%
    addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  
    # draw polygons first...
    addPolygons(data = sde,    fillOpacity = 0.2, weight = 1) %>%
    #addPolygons(data = census_tract_40_west, fillOpacity =  0,color = "black") %>% 
    addPolygons(data = westside,  opacity = 0) %>%
    addPolygons(data = nw1,  opacity = 0.1) %>%
    addPolygons(data = nw2,  opacity = 0.1)%>%
    addPolygons(data = nc1,  opacity = 0.1)%>%
    addPolygons(data = ne1,  opacity = 0.1)%>%
    addPolygons(data = ne2,  opacity = 0.1)%>%
    addPolygons(data = ne3,  opacity = 0.1)%>%
    addPolygons(data = e1,  opacity = 0.1)%>%
    addPolygons(data = sw1,  opacity = 0.1)%>%
    addPolygons(data = sc1,  opacity = 0.1)%>%
    addPolygons(data = sc2,  opacity = 0.1)%>%
    addPolygons(data = sc3,  opacity = 0.1)%>%
    addPolygons(data = se1,  opacity = 0.1)%>%
    addPolygons(data = se2,  opacity = 0.1)%>%
    addPolygons(data = se3,  opacity = 0.1)%>%
    addPolygons(data = se4,  opacity = 0.1)%>%
    addCircleMarkers(
      data        = cha_projects,
      radius      = 1,
      color       = "darkred",
      fillOpacity = 0.6, 
      label       = ~as.character(label),
      labelOptions = labelOptions(
        direction = "auto",
        textsize  = "14px",
        opacity   = 0.9,
        offset    = c(4, -4)
      )
      ) %>%
    addCircleMarkers(
        data        = riots,
        radius      = 1,
        color       = "purple",
        fillOpacity = 0.6, 
        label       = ~as.character(label),
        labelOptions = labelOptions(
          direction = "auto",
          textsize  = "14px",
          opacity   = 0.9,
          offset    = c(4, -4)
        )
        ) %>%
    addControl(paste("Year:", year), position = "topright") %>% 
    setView(lng = center[1], lat = center[2], zoom = 11.5) 
  
  
  for(nb in nhoods) {
    map <- map %>%
      addCircleMarkers(
        data = filter(homs_year, neighborhood == nb),
        radius = 1,
        fillOpacity = 0.7,
        label = ~category,
        group = nb,
        clusterOptions = markerClusterOptions()
      )
  }

   return(map)

  # %>%
  # 
  # # …then draw circles on top
  # addCircleMarkers(
  #   data        = homs_year,
  #   radius      = 1,
  #   color       = "black",
  #   fillOpacity = 0.6,
  #   label       = ~as.character(category),
  #   labelOptions = labelOptions(
  #     direction = "auto",
  #     textsize  = "14px",
  #     opacity   = 0.9,
  #     offset    = c(4, -4)
  #   )
  #   ,
  #   clusterOptions = markerClusterOptions(
  #     showCoverageOnHover = TRUE,
  #     spiderfyOnMaxZoom   = TRUE,
  #     zoomToBoundsOnClick = FALSE,
  #     maxClusterRadius    = 50, 
  #     disableClusteringAtZoom = 15
  #   )
  #   ) 

  
}


```

```{r}
map_year(1940)
map_year(1941)
map_year(1942)
map_year(1943)
map_year(1944)
map_year(1945)
map_year(1946)
map_year(1947)
map_year(1948)
map_year(1949)
map_year(1950)
map_year(1951)
map_year(1952)
map_year(1953)
map_year(1954)
map_year(1955)
map_year(1956)
map_year(1957)
map_year(1958)
map_year(1959)
map_year(1960)
map_year(1961)
map_year(1962)
map_year(1963)
map_year(1964)
map_year(1965)
map_year(1966)
map_year(1967)
map_year(1968)
map_year(1969)
map_year(1970)
map_year(1971)
map_year(1972)
map_year(1973)
map_year(1974)
map_year(1975)
map_year(1976)
map_year(1977)
map_year(1978)
map_year(1979)
map_year(1980)
map_year(1981)
map_year(1982)
map_year(1983)
map_year(1984)
map_year(1985)
map_year(1986)
map_year(1987)
map_year(1988)
map_year(1989)
map_year(1990)
```



#SELECT NEIGHBORHOOD ANALYSIS


```{r}


ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "n"))
       , aes(x= year, y = count, group = neighborhood))  +
  geom_line(aes(color = neighborhood))+
  labs(title = "Central Neighborhoods")

ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "e"))
       , aes(x= year, y = count, group = neighborhood))  +
  geom_line(aes(color = neighborhood))+
  labs(title = "East Neighborhoods")


ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "w"))
       , aes(x= year, y = count, group = neighborhood))  +
  geom_line(aes(color = neighborhood))+
  labs(title = "West Neighborhoods")


ggplot(data = homs_neighborhood_count
       , aes(x= year, y = count, group = neighborhood))  +
  geom_line(aes(color = neighborhood))+
  labs(title = "All Neighborhoods")


```


```{r}

reg <- lm(count ~ as.factor(year) + size, data = homs_neighborhood_count)
summary(reg)
homs_neighborhood_count["residual"]  <- reg$residuals



ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "c"))
       , aes(x= year, y = residual, group = neighborhood))  + 
  geom_line(aes(color = neighborhood)) +
  labs(title = "Central Neighborhoods")


ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "ne"))
       , aes(x= year, y = residual, group = neighborhood))  + 
  geom_line(aes(color = neighborhood))+
  labs(title = "North-East Neighborhoods")


ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "se"))
       , aes(x= year, y = residual, group = neighborhood))  + 
  geom_line(aes(color = neighborhood))+
  labs(title = "South-East Neighborhoods")

ggplot(data = homs_neighborhood_count %>% filter(str_detect(neighborhood, "w"))
       , aes(x= year, y = residual, group = neighborhood))  + 
  geom_line(aes(color = neighborhood))+
  labs(title = "West Neighborhoods")

ggplot(data = homs_neighborhood_count 
       , aes(x= year, y = residual, group = neighborhood))  + 
  geom_line(aes(color = neighborhood))+
  labs(title = "All  Neighborhoods")
```
West neighborhoods seem to be rising by the 70s. We see a central spike in the 60s that continues for a while. Seems to die down in early 2000s but back on the rise. 

SE3 might be interesting - in the late 70s and 80s we seem to be seeing some spikes while all other East neighborhoods are pretty consistently declining. 

# GIFS
## ENTIRE WESTSIDE

```{r}

# Set up the years to iterate through
years <- 1940:2022

year <- year[1]
#even_years <- years[years %% 2 == 0]
center  <- st_coordinates(st_centroid(westside))


# Create and save a map for each year
for (year in years) {
  # Filter data for the current year
  homs_year <- homs_west %>% filter(year == !!year)
  nhoods <- unique(homs_year$neighborhood)
  
  # Calculate standard deviation ellipse for the current year
  sde_data <- std_dev_ellipse(homs_year)
  sde <- st_ellipse(geometry = sde_data, sx = sde_data$sx, sy = sde_data$sy, rotation = -sde_data$theta) 
  
  sde_sf <- st_sf(geometry = sde)%>% 
    mutate(year = !!year)

  
  # Create leaflet map
 map <- leaflet() %>%
    addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
    addPolygons(data = sde,    fillOpacity = 0.2, weight = 1) %>%
    addPolygons(data = westside,  opacity = 0) %>%
    addPolygons(data = nw1,  opacity = 0.1) %>%
    addPolygons(data = nw2,  opacity = 0.1)%>%
    addPolygons(data = nc1,  opacity = 0.1)%>%
    addPolygons(data = ne1,  opacity = 0.1)%>%
    addPolygons(data = ne2,  opacity = 0.1)%>%
    addPolygons(data = ne3,  opacity = 0.1)%>%
    addPolygons(data = e1,  opacity = 0.1)%>%
    addPolygons(data = sw1,  opacity = 0.1)%>%
    addPolygons(data = sc1,  opacity = 0.1)%>%
    addPolygons(data = sc2,  opacity = 0.1)%>%
    addPolygons(data = sc3,  opacity = 0.1)%>%
    addPolygons(data = se1,  opacity = 0.1)%>%
    addPolygons(data = se2,  opacity = 0.1)%>%
    addPolygons(data = se3,  opacity = 0.1)%>%
    addPolygons(data = se4,  opacity = 0.1)%>%
    addCircleMarkers(
      data        = cha_projects,
      radius      = 1,
      color       = "darkred",
      fillOpacity = 0.6, 
      label       = ~as.character(label),
      labelOptions = labelOptions(
        direction = "auto",
        textsize  = "14px",
        opacity   = 0.9,
        offset    = c(4, -4)
      )
      ) %>%
    addCircleMarkers(
        data        = riots,
        radius      = 1,
        color       = "purple",
        fillOpacity = 0.6, 
        label       = ~as.character(label),
        labelOptions = labelOptions(
          direction = "auto",
          textsize  = "14px",
          opacity   = 0.9,
          offset    = c(4, -4)
        )
        ) %>%
    addControl(paste("Year:", year), position = "topright") %>% 
    setView(lng = center[1], lat = center[2], zoom = 11.5) 
  
  
    for(nb in nhoods) {
      map <- map %>%
        addCircleMarkers(
          data = filter(homs_year, neighborhood == nb),
          radius = 1,
          fillOpacity = 0.7,
          label = ~category,
          group = nb,
          clusterOptions = markerClusterOptions()
        )
    }
    
  # Save the map as a PNG image using mapshot
  file_name <- paste0("output/westside/map_images_hotspots/map_", year, ".png")
  mapshot(map, file = file_name)
  # file_name <- paste0("output/westside/map_images_hotspots/map_", year, ".shp")
  # st_write(sde_sf, dsn = file_name,append = FALSE)
  # Print progress
  cat("Created map for year", year, "\n")
}


# Create a GIF from the saved images
# Read all PNG files in the directory
image_files <- list.files("output/westside/map_images_hotspots", pattern = "*.png", full.names = TRUE)
image_files <- sort(image_files)  # Ensure files are in order


# Read images into a list
images <- lapply(image_files, image_read)

# Join images into an animation
animation <- image_join(images)

# Optimize the animation
animation <- image_animate(animation, fps = 2)  # 2 frames per second

# Save as GIF
image_write(animation, "output/westside/homicides_yearly_neighborhoods.gif")
```

```{r}

geojson_files <- list.files("output/westside/map_images", pattern = "*.shp", full.names = TRUE)
geojson_files <- geojson_files %>%
  lapply(read_sf) %>%       # Read each GeoJSON file
  bind_rows()    

all_ellipses <- geojson_files %>%
  sf::st_cast("POLYGON") %>% 
  mutate(size = as.numeric(sf::st_area(geometry))) %>%
  mutate(center = st_centroid(geometry)) %>% 
  mutate(
    prev_center = lag(center),
    # Extract X coordinates (longitude)
    current_x = st_coordinates(center)[,1],
    prev_x = lag(current_x),
    # Calculate distance with sign
    distance_center_prev = ifelse(
      !is.na(prev_center),
      st_distance(center, prev_center, by_element = TRUE) * 
        ifelse(current_x > prev_x, -1, 1),  # Negative for eastward movement
      NA_real_
    )
  ) %>%
  # Remove temporary coordinate columns
  dplyr::select( -prev_x) %>% 
  mutate(decade = floor(year/10)*10)

```

```{r}

ggplot(data = all_ellipses, aes(x = year, y = -current_x)) + 
  geom_line()

```

```{r}

coords_tbl <- st_coordinates(all_ellipses) %>%
  as_tibble() %>%
  rename(lon = X, lat = Y)  

extremes <- coords_tbl %>%
  mutate(year = L2+1939) %>%
  group_by( year) %>%
  summarise(
    east_lon = lon[which.max(lon)][1],
    east_lat = lat[which.max(lon)][1],
    west_lon = lon[which.min(lon)][1],
    west_lat = lat[which.min(lon)][1],
    .groups = "drop"
  ) %>% 
  ungroup() %>% 
  mutate(
    long_distance_m = distHaversine(
      cbind(west_lon, west_lat),
      cbind(east_lon, east_lat)
    )
  )



```

```{r}

ggplot(data = extremes) + 
  geom_line( aes(x = year, y = -east_lon))+ 
  geom_line( aes(x = year, y = -west_lon))+
  geom_smooth(aes(x = year, y = -west_lon))+
  geom_smooth(aes(x = year, y = -east_lon))


ggplot() + 
  geom_point(data = extremes, aes(x = year, y = long_distance_m)) +
  geom_smooth(data = extremes, aes(x = year, y = long_distance_m)) +
  geom_point(data = homs_west %>% st_drop_geometry() %>% group_by(year) %>% summarise(total = n()),
             aes(x = year, y = total*30), color = "orange") +
   scale_y_continuous(
    name      = "East–West Distance (m)",
    limits    = c(0, 10000),        # match your real distance range
    sec.axis  = sec_axis(
      ~ . / 30,                       # undo the *30
      name = "Homs West Count"
    )
  ) 


```
```{r}
counts <- homs_west %>% st_drop_geometry() %>% group_by(year) %>% summarise(total = n())


extremes_count <- extremes %>% 
  left_join(counts)
reg <- lm(long_distance_m ~ total, data =  extremes_count)
extremes_count["residual_distance"] <- reg$residuals


ggplot() + 
  geom_point(data = extremes_count, aes(x = year, y = residual_distance)) +
  geom_smooth(data = extremes_count, aes(x = year, y = residual_distance)) 
```
## NEIGHBORHOODS

```{r}

neighborhood <- "sc1"

neighborhood_map <- read_sf(paste0("data/mst/neighborhoods/", neighborhood, ".shp"))

# Set up the years to iterate through
years <- 1940:2022

year <- years[1]
#even_years <- years[years %% 2 == 0]
center  <- st_coordinates(st_centroid(neighborhood_map))

homs_select <- homs_all %>% 
  st_join(neighborhood_map, join = st_intersects, left = FALSE)
# Create and save a map for each year
for (year in years) {
  # Filter data for the current year
  homs_year <- homs_select %>% filter(year == !!year)
  
  
  # Calculate standard deviation ellipse for the current year
  # sde_data <- std_dev_ellipse(homs_year)
  # sde <- st_ellipse(geometry = sde_data, sx = sde_data$sx, sy = sde_data$sy, rotation = -sde_data$theta) 
  # 
  # sde_sf <- st_sf(geometry = sde)%>% 
  #   mutate(year = !!year)
  # 
  # 
  # Create leaflet map
  
  if (nrow(homs_year) == 0) {
     map <- leaflet() %>%
       addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
       addPolygons(data = neighborhood_map,  opacity = 0) %>%
       addControl(paste("Year:", year), position = "topright") %>% 
       setView(lng = center[1], lat = center[2], zoom = 13) 
    
  } else {
    map <- leaflet() %>%
      addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
      #addPolygons(data = sde,    fillOpacity = 0.2, weight = 1) %>%
      addPolygons(data = neighborhood_map,  opacity = 0) %>%
      addCircleMarkers(
        data        = cha_projects,
        radius      = 1,
        color       = "darkred",
        fillOpacity = 0.6, 
        label       = ~as.character(label),
        labelOptions = labelOptions(
          direction = "auto", 
          textsize  = "14px",
          opacity   = 0.9,
          offset    = c(4, -4)
          )
        ) %>%
      addCircleMarkers(
          data        = riots,
          radius      = 1,
          color       = "purple",
          fillOpacity = 0.6, 
          label       = ~as.character(label),
          labelOptions = labelOptions(
            direction = "auto",
            textsize  = "14px",
            opacity   = 0.9,
            offset    = c(4, -4)
          )
          ) %>%

      addCircleMarkers(
        data        = homs_year,
        radius      = 1,
        color       = "black",
        fillOpacity = 0.6,
        label       = ~as.character(category),
        labelOptions = labelOptions(
          direction = "auto",
          textsize  = "14px",
          opacity   = 0.9,
          offset    = c(4, -4)
        )
        ,
        clusterOptions = markerClusterOptions(
          showCoverageOnHover = TRUE,
          spiderfyOnMaxZoom   = TRUE,
          zoomToBoundsOnClick = FALSE,
          maxClusterRadius    = 50, 
          disableClusteringAtZoom = 15
        )
      ) %>%
  addControl(paste("Year:", year), position = "topright") %>% 
  setView(lng = center[1], lat = center[2], zoom = 13) 
    
  }

  
  # Save the map as a PNG image using mapshot
  file_name <- paste0("output/westside/map_images_hotspots/map_", year, ".png")
  mapshot(map, file = file_name)
  #file_name <- paste0("output/westside/map_images_hotspots/map_", year, ".shp")
  #st_write(sde_sf, dsn = file_name,append = FALSE)
  # Print progress
  cat("Created map for year", year, "\n")
}

# Create a GIF from the saved images
# Read all PNG files in the directory
image_files <- list.files("output/westside/map_images_hotspots", pattern = "*.png", full.names = TRUE)
image_files <- sort(image_files)  # Ensure files are in order




# Read images into a list
images <- lapply(image_files, image_read)

# Join images into an animation
animation <- image_join(images)

# Optimize the animation
animation <- image_animate(animation, fps = 2)  # 2 frames per second

# Save as GIF
image_write(animation, paste0("output/westside/homicides_yearly_", neighborhood, ".gif"))
```


Split west side into different hot spot areas 
Show how the count varies over time 




Racine - Halstead - Roosevelt 
- 1940: 5 
- 1941: 11 
- 1942: not a big crime year, 2 
- 1943: 8
- 1944: 7
- 1945: 3 
- 1946: 11
- 1947: 9
- 1948: ~ 11 but some a little westward


Along Roosevelt 

- 1944: 12
- 1945: 0
- 1946: 4
- 1947: 10-15


Blocks West of Union Park

- 1946: 12
- 1947: 4
- 1948: ~ 11-12

Along Kennedy Expressway

1947: 10 (burns - seems to be fire)


What's happening from 43 to 44


